Bayesian Nonparametric Neural Image Segmentation

### 2013-01-31 latex/submission.tex
most substance is done

\documentclass[draft,english]{article}
\usepackage{fullpage}
\usepackage{amsmath}
\usepackage{tikz}
\usepackage{tkz-base}
\usetikzlibrary{shapes,decorations,shadows}
\usetikzlibrary{decorations.pathmorphing}
\usetikzlibrary{decorations.shapes}
\usetikzlibrary{fadings}
\usetikzlibrary{patterns}
\usetikzlibrary{calc}
\usetikzlibrary{decorations.text}
\usetikzlibrary{decorations.footprints}
\usetikzlibrary{decorations.fractals}
\usetikzlibrary{shapes.gates.logic.IEC}
\usetikzlibrary{shapes.gates.logic.US}
\usetikzlibrary{fit,chains}
\usetikzlibrary{positioning}
\usepgflibrary{shapes}
\usetikzlibrary{scopes}
\usetikzlibrary{arrows}
\usetikzlibrary{backgrounds}
\tikzset{latent/.style={circle,fill=white,draw=black,inner sep=1pt, 
minimum size=20pt, font=\fontsize{10}{10}\selectfont},
obs/.style={latent,fill=gray!33},
const/.style={rectangle, inner sep=1pt},
factor/.style={rectangle, fill=black,minimum size=5pt, inner sep=1pt},
>={triangle 45}}

@@

\title{A Bayesian Nonparametric Approach to Neural Image Segmentation}
\author{Kui Tang and Frank Wood}
\date{\today}
\maketitle
\begin{abstract}
We develop a Bayesian nonparametric method to segment neurons in electron micrograph images of brain tissue. Our model imposes spatial locality on arbitrary features, automatically determines the number of cells in an image, and can be used in both unsupervised and semi-supervised settings. We model the distribution of neurons in an image as an infinite Gaussian mixture model. Each neuron's pixels are characterized by likelihood function that is a combination of multivariate Gaussian and naive Bayes features. Using data collected by Kasthuri et al., we show that our model can obtain plausible segmentations with no training data, while receiving a significant boost in its presence.

\section{Introduction}
A new branch of computational neuroscience, \emph{connectomics}, aims to reconstruct the physical location and circuit connectivity of the brain. Recent advances in electron microscopy have produced images of 10 nm in the lateral plane and 50 nm in the vertical plane, sufficient detail to capture most of even the smallest neural processes~\cite{Briggman2006}. The scale of the data is profound; the human brain is estimated to contain 100 billion neurons connected by up to a quadrillion synapses~\cite{Kasthuri2010}. Like the human genome project before, it is clear that automated methods are needed to map the connectome.

All published connectomes have been manually labelled. White et al. published the first connectome of \emph{C. elegans} in 1986, marking axons, dendrites, and cell bodies in each frame and tracing the labels through frames~\cite{White12111986}. More recent work uses computer assistance but fundamentally rely on a skilled human's identification of neural bodies~\cite{Jarrell2012,Bock2011Roberts2011}. Several authors have developed efficient semi-automated methods \cite{Roberts2011,Unger2009}. Unfortunately, current baseline speeds of computer-assisted manual tracing would still take 10,000 person-years to construct a mouse cortical column~\cite{Briggman2006}. Therefore, henceforth we focus on automated techniques.

\section{Related Work}
\fxfatal{Write topic sentence, summarize your current literature review, and add paragraphs about feature engineering and graph pathfinding methods. Cite Malik's normalized cuts.}
Automated techniques generally begin by segmenting a 2D slice based on cell membrane boundaries---a highly distinctive feature.

@@

\fxfatal{What microscopy method did they use? and put in the right place.}
We analyzed a  volume of mouse visual cortex of resolution $3 \times 3 \times 30 \mbox{ nm}^3$ collected by \fxfatal{Unpublished; to cite}{Kasthuri et al.}

@@

\begin{tikzpicture}
\matrix[row sep=6mm, column sep=8mm, matrix anchor=mid] (mdp) {
  \node (alpha) [const] {$\alpha$}; & & \\
  \node (pi) [latent] {$\+\pi$}; & \node (beta) [latent] {$\beta_k$}; & \node (alphabeta) [const] {$\alpha_0, \beta_0$}; \\
  \node (z) [latent] {$z_n$}; & \node(Lambda) [latent] {$\Lambda_k$}; & \node(Lambdanu) [const] {$\Lambda_0, \nu$}; \\
  \node (x) [obs] {$\+x_n$}; & \node(mu) [latent] {$\+\mu_k$}; & \node(mukappa) [const] {$\+\mu_0, \kappa$}; \\
};

@@

\end{tikzpicture}
\caption{The graphical model} \label{fig:model}
\end{figure}

\label{sec:model}
We present our model in Figure~\ref{fig:model}. Below, we explain our feature representation and how we overcome the aforementioned difficulties. First, the number of cells must be inferred from the image itself. To do so, we employ a Mixture Dirichlet process, discussed in section~\ref{sec:mdp}. Further, the visual complexity and noise of electron micrograph images do not yield the clear separations required for assumptions 2 and 3. We fix this with Radon-like features to separate cell interiors and boundaries in section~\ref{sec:radon}, and constrain cluster assignments with a geodesic distance that penalizes boundary crossings in section~\ref{sec:geodesic}.

Pixel $i$ has feature vector $\+{x}_n = (x_n, y_n, r_n, f_n, \+{d}_n)$ where $x_n, y_n$ represent the pixel's spatial coordinates, $r_n \in [0,1]$ is the raw pixel intensity, $f_n \in [0,1]$ is the average response of the interior Radon-like feature, and $\+{d}_n \in \+{R}^N$ is a vector of geodesic distances to every other pixel. If position $\+p \in \+R^2$, when discretized, is the location of pixel $n$ and the discrete $\+q$ is the location of pixel $m$, then we write $d(\+p, \+q) = \+d_n,m$. Putting these components together yields the model in Figure~\ref{fig:model}. The generative story is:

\begin{itemize}
\item $\+\pi \sim \mbox{DirichletProcess}(\alpha)$
\item $k | \+\pi \sim \mbox{Discrete}(\+\pi)$
\item If $k$ is a new (empty) cluster:
  \begin{itemize}
  \item $\Lambda_k | T, \nu \sim \mbox{Wishart}(\Lambda | T, \nu)$ (in the notation of \cite{Murphy2003})
  \item $\+\mu_k | \kappa , \Lambda_k \sim \mathcal{N}\left( \+\mu_k | \+\mu_0 , (\kappa\Lambda)^{-1} \right)$
  \item $\beta_k | \alpha_0, \beta_0 \sim \mbox{Gamma}(\beta_k | \alpha_0, \mbox{rate} = \beta_0)$
  \end{itemize}
\item $\+x | \+\mu_k , \Lambda_k, \beta_k \propto \mathcal{N}(\+x_{1:4} | \+\mu_k , \Lambda_k) \mbox{Exponential}\left( d(\+x_{1:2}, \+\mu_{k,1:2}) | \mbox{rate} = \beta_k \right)$ (in the notation of section~\ref{sec:geodesic})
\end{itemize}

@@

The indicators $z_n$ are drawn from the discrete distribution $\+\pi$, a $K$-dimensional vector which is in turn is drawn from a symmetric Dirichlet distribution parameterized by $\alpha$: $$p \left( \+\pi | \alpha, K \right) = \frac{ \Gamma ( \alpha ) }{ \Gamma ( \alpha/K ) ^K} \prod_{k=1}^K \+\pi_k^{\alpha/K - 1}.$$ Because the Dirichlet and discrete distributions are conjugate, we can marginalize $\+\pi$. Let $\+{z}_{-n}$ denote the vector of all indicators excluding element $n$ and $N_{k,-n}$ denote the number of points assigned to class $k$ excluding point $n$. A standard result \fxfatal{CITATION} yields the posterior $$p(z_n = k | \+{z}_{-n})  = \frac{ N_{k,-n} + \alpha/K }{ N - 1 + \alpha }$$ which makes Gibbs sampling easy.

Now consider $K \rightarrow \infty$. First, the term $\alpha / K \rightarrow 0$. Given finitely many observations, we have only finitely many components $k$ with $N_{k,-n} > 0$. Therefore, the density 
\[
p(z_{n}=k|\+{z}_{-n},\alpha)=\begin{cases}
\frac{N_{k,-n}}{N-1+\alpha} & N_{k,-n}>0\\
\frac{\alpha}{N-1+\alpha} & \mbox{otherwise}
\end{cases}
\]
satisfies the limiting conditions and correctly normalizes to 1 when summed with $k = 1, \ldots, \infty$. We now say that the mixing distribution $\+\pi$, which is now infinite-dimensional, is drawn from a \emph{Dirichlet process}. Since conjugacy allows us to eliminate $\+\pi$ from computations, Gibbs sampling from an infinite mixture is just as efficient as for a finite mixture case.

\subsection{Radon-like Features}
\label{sec:radon}
Ideally, for a mixture model, pixel intensities within a cell should be homogeneous. But the diversity of intracellular structures and significant imaging noise yield data far from this ideal. As a countermeasure, we augment raw pixel intensities with $\emph{Radon-like feature}$ responses, which have proven useful for connectomics segmentation \cite{Kumar2010}. \fxfatal{Make this clearer... Even the original paper doesn't have a concise description}. A Radon-like feature aggregates a function of image intensity over regions defined by structural constraints, such as edges. Each feature is measured from a single direction, which we can represent as a \emph{scan line} $\+{l}(t) : [0, 1] \rightarrow \+{R}^2$. We represent the structure information as a set of points $\{ t_1, \ldots, t_m \}$ along line $\+{l}$. We follow the practice in \cite{Kumar2010} to evaluate Radon-like features along uniform angular intervals (we find 72 to avoid most discretization artifacts) and take the mean response from all directions. Of course, using the entire vector of feature responses with a multivariate likelihood is also possible. Moreover, in \cite{Kumar2010} as well as this paper, we will use the filter response $$R(x,y) = \max_{\sigma, \phi} \Delta G(\sigma, \phi) \ast I(x,y) $$ where $\Delta G(\sigma, \phi)$ is the Gaussian second derivative filter at scale $\sigma$ and orientation $\phi$. We use $R$ to (1) obtain knots, by passing to a Canny edge mapper, and (2) use as input to extraction functions (discussed below). Unlike other methods, we use the noisy edge response only as a guide for additional feature extraction, not as a testimony for the cell segments themselves.

Let $T(I, x, y)$ be an $\emph{extraction function}$ which maps an image and $xy$-coordinates to a scalar. Suppose a line $\+{l}$ is given and $(x, y) = \+{l}(t)$ for some $t \in [t_i, t_{i+1}]$. Then the Radon-like response for image $I$ at $(x, y)$ in direction $\+{l}$ is given by $$\Psi \left( I,\+{l},t \right) = T \left( I, \+{l}(t) \right).$$ We will use the extraction function for cell background segmentation $$T_{\mbox{bg}}(I, \+{l}(t)) = \min_{t \in [t_i, t_{i+1}]} \left\{ I \left( \+{l}(t) \right) \right\}$$ which simply returns the dimmest pixel between any two knots \cite{Kumar2010}. Because knots denote edges, this feature tends to homogenize pixel intensities within a cell.

Finally, we want pixels to not cross cell boundaries. The first step in encoding this constraint is to use a boundary-enhancing extraction function \cite{Kumar2010} $$T_{\mbox{bd}} = \frac{\int_t^{t+1} R\left( \+{l}(\tau) \right) d\tau }{\left\Vert \+{l}(t_{i+1}) - \+{l}(t_{i}) \right\Vert_2 }.$$ This function assigns to each pixel the average value of the Gaussian second-derivative response. Recall that the GSD tends to enhance boundaries, the most salient of which are captured by the Canny detector as knots. Thus, $T_{\mbox{bd}}$ smoothes the spurious GSD responses, avoiding oversegmentation. \fxfatal{Tighten up wording here}. Further, in connectomics data, cell membranes tend to form thick boundaries. But classic edge detectors work best at single-pixel gradients and return two edges for a thick boundary. \fxfatal{cite Rafael}. In particular, the Canny detector produces two knots on either side of a cell boundary. In the GSD response $R$, cell boundaries are consistently bright, so the integral in $T_{\mbox{bd}}$ returns a distinctly brighter response on boundaries, regardless of thickness. We use this boundary enhancement information in the next section.

\subsection{Geodesic Distance}
\label{sec:geodesic}
Figure~\ref{fig:radon_bd} illustrates that cells need not be elliptical, or even convex. All that can be said is that they are contained in a closed loop that is generally gives a distinct Radon-like boundary response. Therefore, shape constraints based on Euclidean distances are not justified. Instead, we want to encode the a priori belief that "the cell body should not spill outside its membrane." We quantify this property with a \emph{geodesic distance}. Construct a graph whose vertices are the pixels and undirected edge for adjacent pixels ($(x,y)$ is an edge if $y$ is in the 1-pixel neighborhood of $x$). Each edge has a weight $w(i,j) = \left| b_i - b_j \right|$ where $b_k$ is the boundary Radon-like response at pixel $k$. We define $d(\+{x}_i, \+{x}_j)$ to be the total weight of the shortest path between pixels $i$ and $j$, which is computed with Johnson's algorithm. In practice, the $O(N^2)$ storage is prohibitive, so we instead we downsample the image into SLIC superpixels \cite{Achanta2012} and compute superpixel-based distances.Address practical concerns. Further, because individual cells are tiny relative to an entire slice, only local distances are needed, because far points must be outside the current cell. All paths crossing a cell boundary must traverse pixels with low, high, and again low boundary feature responses. Thus, the shortest path wil also be one with a high weight. So if pixels $i$ and $j$ do not belong to the same cell, defined by the region enclosed by a loop of pixels with high boundary feature response, then $d(\+{x}_i, \+{x}_j)$ is high.

@@

\section{Inference}
Our model encodes spatial locality in two novel ways: in the Gaussian covariance which bias class membership to Euclidean-close points, and in the geodesic distances which penalizes boundary crossings.

>>>>

\documentclass[english]{article}
\usepackage{fullpage}
\usepackage{amsmath}
\usepackage{tikz}
\usepackage{tkz-base}
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\graphicspath{{../fig/}}

\usetikzlibrary{shapes,decorations,shadows}
\usetikzlibrary{decorations.pathmorphing}
\usetikzlibrary{decorations.shapes}
\usetikzlibrary{fadings}
\usetikzlibrary{patterns}
\usetikzlibrary{calc}
\usetikzlibrary{decorations.text}
\usetikzlibrary{decorations.footprints}
\usetikzlibrary{decorations.fractals}
\usetikzlibrary{shapes.gates.logic.IEC}
\usetikzlibrary{shapes.gates.logic.US}
\usetikzlibrary{fit,chains}
\usetikzlibrary{positioning}
\usepgflibrary{shapes}
\usetikzlibrary{scopes}
\usetikzlibrary{arrows}
\usetikzlibrary{backgrounds}
\tikzset{latent/.style={circle,fill=white,draw=black,inner sep=1pt, 
minimum size=20pt, font=\fontsize{10}{10}\selectfont},
obs/.style={latent,fill=gray!33},
const/.style={rectangle, inner sep=1pt},
factor/.style={rectangle, fill=black,minimum size=5pt, inner sep=1pt},
>={triangle 45}}

@@

\title{Bayesian Nonparametric Neural Image Segmentation}
\author{Kui Tang and Frank Wood}
\date{\today}
\maketitle
\begin{abstract}
We develop a Bayesian nonparametric method to segment neurons in electron micrograph images of brain tissue. Our model imposes spatial locality on arbitrary features, automatically determines the number of cells in an image, and can be used in both unsupervised and semi-supervised settings. We introduce a novel geodesic distance which adaptively adjust segmentation to respect detected cell boundaries. Using data collected by Kasthuri et al., we show that our model can obtain plausible segmentations with no training data, while receiving a significant boost in its presence. Finally, the Bayesian nonparametric model offers a natural extension to 3 dimensions, where neural columns of segments \emph{through} slices are modelled as coagulation-fragmentation processes, paving the road for a fully Bayesian solution to the connectomics problem.
\end{abstract}

\section{Introduction}
A new branch of computational neuroscience, \emph{connectomics}, aims to reconstruct the physical location and circuit connectivity of the brain. Recent advances in electron microscopy have produced images of 10 nm in the lateral plane and 50 nm in the vertical plane, sufficient detail to capture most of even the smallest neural processes~\cite{Briggman2006}. The scale of the data is profound; the human brain is estimated to contain 100 billion neurons connected by up to a quadrillion synapses~\cite{Kasthuri2010}.

All published connectomes have been manually labelled. White et al. published the first connectome of \emph{C. elegans} in 1986, marking axons, dendrites, and cell bodies in each frame and tracing the labels through frames~\cite{White12111986}. Recent work in semi-automated methods are accurate and promising \cite{Roberts2011,Unger2009,Jarrell2012,Bock2011}, but current baseline speeds of computer-assisted manual tracing would still take 10,000 person-years to construct a mouse cortical column~\cite{Briggman2006}, hence the need for fully automatic methods.

\section{Related Work}
Automated techniques generally begin by finding the cell boundaries---a highly distinctive feature. Because cells are usually disconnected, segmentation can be reduced to binary classification of whether voxels are inside or outside a cell. Segmentation can be done directlyin voxel space. Jain et al. \cite{Jain2007} achieved an 8.0\% test error with a 34,000 convolutional neural network, a statistically significant improvement other the 18.95\% error of thresholding, while MRFs and CRFs do no better than thresholding. Andres et al. obtain similar performance, at 8\% misclassification, using the watershed algorithm over a probability map of a random forest with hand-engineered features \cite{Andres2008}. However, the dataset used by both used a special intracellular stain which made segmentation easier. The generality of these solutions are thus unclear.

More common is to segment in pixel space and track the identity of segmented bodies in the vertical dimension. This mitigates the poor vertical resolution of EM images \cite{Kaynig2010, Briggman2006}. Jurrus et al. \cite{Jurrus2008} first obtain 2D segmentations with watershed after preprocessing linear diffusion filtering (NB: learn what this is!), and construct a weighted graph of these segments, with edges between adjacent sections with edge costs favoring high cross-correlation of segments low lateral displacement. Each neural process is thus a shortest path through this graph, computed by Dijkstra's algorithm. This cost function is robust to merging and loss of sections, but because Dijkstra's algorithm finds a single path, it cannot model branching. Kaynig et al. \cite{Kaynig2010a} presuppose segmentation and represent the image as a graph of regions, like Jurrus. Pairwise features are represented as edge weights, and a random forest classifier is trained to output a scalar similarity score for each pair. Then a probabilistic model encoding geometric constraints with cross-section correspondences as hidden variables is trained with EM. Finally, the maximum-likelihood result is the 3D reconstruction. These pairwise correspondences are fed to an agglomerative clustering algorithm, constructing the neural processes from ground-up in the same manner as a human. Unlike in Jurrus, processes may merge. However, each region is explicitly limited to two correspondences: one above and below, so processes still cannot branch.

Feature engineering significantly shapes the state of the art. Venkataraju et al. \cite{Venkataraju2009} generated 100 Hessian neighborhood features at each pixel and classified Jurrus's dataset with AdaBoost. Lucchi et al. \cite{Lucchi2010} develop ray descriptors, which capture shape features without a predefined shape model. These features are rotation, scale, and translation invariant, but strongly discriminate the types of objects observed in EM images. There features were used in SVMs to classify interior and boundary points of mitochondria, which are considerably harder to classify than other cellular structures \cite{Kaynig2010b}, and a graph cut found the final 2D segmentation with 98\% accuracy. Kaynig et al. encode perceptual grouping constraints---rules of thumb used by human experts to group neuron structures---as pairwise interaction terms in an energy function solved by graph cut. Minimizing the energy function gives a segmentation, and identifying maximum overlapping regions between slices is sufficient to obtain a 3D reconstruction

\fxwarning{Who cites Malik's normalized cuts?}

@@

\fxwarning{What microscopy method did they use? and put in the right place.}
We analyzed a  volume of mouse visual cortex of resolution $3 \times 3 \times 30 \text{ nm}^3$ collected by \fxwarning{Unpublished; to cite}{Kasthuri et al.}

@@

\begin{tikzpicture}
\matrix[row sep=6mm, column sep=8mm, matrix anchor=mid] (mdp) {
  \node (alpha) [const] {$\alpha$}; & & \\
  \node (pi) [latent] {$\+\pi$}; & \node (beta) [latent] {$\gamma_k$}; & \node (alphabeta) [const] {$\beta_0, \gamma_0$}; \\
  \node (z) [latent] {$z_n$}; & \node(Lambda) [latent] {$\Lambda_k$}; & \node(Lambdanu) [const] {$\Lambda_0, \nu$}; \\
  \node (x) [obs] {$\+x_n$}; & \node(mu) [latent] {$\+\mu_k$}; & \node(mukappa) [const] {$\+\mu_0, \kappa$}; \\
};

@@

\end{tikzpicture}
\caption{The graphical model. Unlike a standard Gaussian mixture model, our model couples $\gamma_k$ and $\+\mu_k$ through the $\+x_n$, giving the flexibility to model boundary constraints, but also presents challenges in inference. See text for details.}\label{fig:model}
\end{figure}

\label{sec:model}
We present our model in Figure~\ref{fig:model}. Below, we explain our feature representation and how we overcome the aforementioned difficulties. First, the number of cells must be inferred from the image itself. To do so, we employ a mixture Dirichlet process, discussed in section~\ref{sec:mdp}. Further, the visual complexity and noise of electron micrograph images do not yield the clear separations required for assumptions 2 and 3. We fix this with Radon-like features to separate cell interiors and boundaries in section~\ref{sec:radon}. Finally, our model encodes spatial locality in two novel ways: in the Gaussian covariance which bias class membership to Euclidean-close points, and in the geodesic distances which penalizes boundary crossings, discussed in section~\ref{sec:geodesic}.

Pixel $i$ has feature vector $\+{x}_n = (x_n, y_n, r_n, f_n, \+{d}_n)$ where $x_n, y_n$ represent the pixel's spatial coordinates, $r_n \in [0,1]$ is the raw pixel intensity, $f_n \in [0,1]$ is the average response of the interior Radon-like feature, and $\+{d}_n \in \+{R}^N$ is a vector of geodesic distances to every other pixel. If position $\+p \in \+R^2$, when discretized, is the location of pixel $n$ and the discrete $\+q$ is the location of pixel $m$, then we write $d(\+p, \+q) = \+d_{n,m}$. Putting these components together yields the model in Figure~\ref{fig:model}. The generative story is:

\begin{itemize}
\item $\+\pi \sim \text{DirichletProcess}(\alpha)$
\item $k | \+\pi \sim \text{Discrete}(\+\pi)$
\item If $k$ is a new (empty) cluster:
  \begin{itemize}
  \item $\Lambda_k | \Lambda_0, \nu \sim \text{Wishart}(\Lambda | \Lambda_0, \nu)$ (in the notation of \cite{Murphy2003})
  \item $\+\mu_k | \kappa , \Lambda_k \sim \mathcal{N}\left( \+\mu_k | \+\mu_0 , (\kappa\Lambda)^{-1} \right)$
  \item $\gamma_k | \beta_0, \gamma_0 \sim \text{Gamma}(\gamma_k | \beta_0, \text{rate} = \gamma_0)$
  \end{itemize}
\item $\+x | \+\mu_k , \Lambda_k, \gamma_k \propto \mathcal{N}(\+x_{1:4} | \+\mu_k , \Lambda_k) \text{Exponential}\left( d(\+x_{1:2}, \+\mu_{k,1:2}) | \text{rate} = \gamma_k \right) \label{eq:xlike}$ (in the notation of section~\ref{sec:geodesic})
\end{itemize}

@@

The indicators $z_n$ are drawn from the discrete distribution $\+\pi$, a $K$-dimensional vector which is in turn is drawn from a symmetric Dirichlet distribution parameterized by $\alpha$: $$p \left( \+\pi | \alpha, K \right) = \frac{ \Gamma ( \alpha ) }{ \Gamma ( \alpha/K ) ^K} \prod_{k=1}^K \+\pi_k^{\alpha/K - 1}.$$ Because the Dirichlet and discrete distributions are conjugate, we can marginalize $\+\pi$. Let $\+{z}_{-n}$ denote the vector of all indicators excluding element $n$ and $N_{k,-n}$ denote the number of points assigned to class $k$ excluding point $n$. A standard result \cite{Rasmussen2000} yields the posterior $$p(z_n = k | \+{z}_{-n})  = \frac{ N_{k,-n} + \alpha/K }{ N - 1 + \alpha }$$ which makes Gibbs sampling easy.

Now consider $K \rightarrow \infty$. First, the term $\alpha / K \rightarrow 0$. Given finitely many observations, we have only finitely many components $k$ with $N_{k,-n} > 0$. Therefore, the density 
\[
p(z_{n}=k|\+{z}_{-n},\alpha)=\begin{cases}
\frac{N_{k,-n}}{N-1+\alpha} & N_{k,-n}>0\\
\frac{\alpha}{N-1+\alpha} & \text{otherwise}
\end{cases} \label{eq:mdp}
\]
satisfies the limiting conditions and correctly normalizes to 1 when summed over $k = 1, \ldots, \infty$. We now say that the mixing distribution $\+\pi$, which is now infinite-dimensional, is drawn from a \emph{Dirichlet process}. Since conjugacy allows us to eliminate $\+\pi$ from computations, Gibbs sampling from an infinite mixture is just as efficient as for a finite mixture case.

\subsection{Radon-like Features}
\label{sec:radon}

\begin{figure}
\centering
\includegraphics[width=0.32\linewidth, clip]{JAN_im}
\includegraphics[width=0.32\linewidth, clip]{JAN_radon_bg}
\includegraphics[width=0.32\linewidth, clip]{JAN_radon_bd}
\caption{A $600\times600$ image slice, its background average Radon response, and its boundary average Radon response. The background response homogenizes pixel intensities within a cell body, while the boundary response highlights cell boundaries.} \label{fig:radon}
\end{figure}

Ideally, for a mixture model, pixel intensities within a cell should be homogeneous. But the diversity of intracellular structures and significant imaging noise yield data far from this ideal. To correct this, we augment raw pixel intensities with $\emph{Radon-like feature}$ responses, which have proven useful for connectomics segmentation \cite{Kumar2010}. \fxwarning{Make this clearer... Even the original paper doesn't have a concise description}. A Radon-like feature aggregates a function of image intensity over regions defined by structural constraints, such as edges. Each feature is measured from a single direction, which we can represent as a \emph{scan line} $\+{l}(t) : [0, 1] \rightarrow \+{R}^2$. We represent the structure information as a set of points $\{ t_1, \ldots, t_m \}$ along line $\+{l}$. We follow the practice in \cite{Kumar2010} to evaluate Radon-like features along uniform angular intervals (we find 72 to avoid most discretization artifacts) and take the mean response from all directions. Of course, using the entire vector of feature responses with a multivariate likelihood is also possible. Moreover, in \cite{Kumar2010} as well as this paper, we will use the filter response $$R(x,y) = \max_{\sigma, \phi} \Delta G(\sigma, \phi) \ast I(x,y) $$ where $\Delta G(\sigma, \phi)$ is the Gaussian second derivative filter at scale $\sigma$ and orientation $\phi$. We use $R$ to (1) obtain knots, by passing to a Canny edge mapper, and (2) use as input to extraction functions (discussed below). Unlike other methods, we use the noisy edge response only as a guide for additional feature extraction, not as a testimony for the cell segments themselves.

Let $T(I, x, y)$ be an $\emph{extraction function}$ which maps an image and $xy$-coordinates to a scalar. Suppose a line $\+{l}$ is given and $(x, y) = \+{l}(t)$ for some $t \in [t_i, t_{i+1}]$. Then the Radon-like response for image $I$ at $(x, y)$ in direction $\+{l}$ is given by $$\Psi \left( I,\+{l},t \right) = T \left( I, \+{l}(t) \right).$$ We will use the extraction function proposed in \cite{Kumar2010} for cell background segmentation $$T_{\text{bg}}(I, \+{l}(t)) = \min_{t \in [t_i, t_{i+1}]} \left\{ I \left( \+{l}(t) \right) \right\}$$ which simply returns the dimmest pixel between any two knots. Because knots denote edges, this feature tends to homogenize pixel intensities within a cell.

Finally, we want pixels to not cross cell boundaries. The first step in encoding this constraint is to use a boundary-enhancing extraction function \cite{Kumar2010} $$T_{\text{bd}} = \frac{\int_t^{t+1} R\left( \+{l}(\tau) \right) d\tau }{\left\Vert \+{l}(t_{i+1}) - \+{l}(t_{i}) \right\Vert_2 }.$$ This function assigns to each pixel the average value of the Gaussian second-derivative response. Recall that the GSD tends to enhance boundaries, the most salient of which are captured by the Canny detector as knots. Thus, $T_{\text{bd}}$ smoothes the spurious GSD responses, avoiding oversegmentation. \fxwarning{Tighten up wording here.}Further, in connectomics data, cell membranes tend to form thick boundaries. But classic edge detectors work best at single-pixel gradients and return two edges for a thick boundary. In particular, the Canny detector produces two knots on either side of a cell boundary. In the GSD response $R$, cell boundaries are consistently bright, so the integral in $T_{\text{bd}}$ returns a distinctly brighter response on boundaries, regardless of thickness. We use this boundary enhancement information in the next section.

\subsection{Geodesic Distance}
\label{sec:geodesic}
\begin{figure}
\includegraphics[width=0.24\linewidth, clip]{JAN_bd_small}
\includegraphics[width=0.24\linewidth, clip]{JAN_bd_dist}
\includegraphics[width=0.24\linewidth, clip]{JAN_bd_thresh}
\includegraphics[width=0.24\linewidth, clip]{JAN_bd_hist}
\caption{Boundary Radon-like feature response; red marks the center from which the remaining pictures are measured. Grayscale map of geodesic distance (of superpixels). Threshold of geodesic distance $< 0.1$. Distribution of geodesic distances about the red center. Although the separation between interior and exterior is easy effort, BASELINE BLAH}\label{fig:geodesic}
\end{figure}
Figures~\ref{fig:radon} and~\ref{fig:geodesic} illustrates that cells need not be elliptical, or even convex. All that can be said is that they are contained in a closed loop that is generally gives a distinct Radon-like boundary response. Therefore, shape constraints based on Euclidean distances are not justified. Instead, we want to encode the a priori belief that "the cell body should not spill outside its membrane." Let us quantify this statement. Construct an undirected graph with pixels as vertices and pairs of adjacent pixels as edges ($(x,y)$ is an edge if $y$ is in the 1-pixel neighborhood of $x$). Each edge has a weight $w(i,j) = \left| b_i - b_j \right|$ where $b_k$ is the boundary Radon-like response at pixel $k$. We define the \emph{geodesic distance} $d(\+{x}_i, \+{x}_j)$ to be the total weight of the shortest path between pixels $i$ and $j$, which is computed with Johnson's algorithm. In practice, the $O(N^2)$ storage is prohibitive, so we instead we downsample the image into SLIC superpixels \cite{Achanta2012} and compute superpixel-based distances. Further, because individual cells are tiny relative to an entire slice, only local distances are needed, because far points must be outside the current cell. All paths crossing a cell boundary must traverse pixels with low, high, and again low boundary feature responses. Thus, the shortest path wil also be one with a high weight. So if pixels $i$ and $j$ do not belong to the same cell, defined by the region enclosed by a loop of pixels with high boundary feature response, then $d(\+{x}_i, \+{x}_j)$ is high.

@@

\section{Inference}
Inference is performed by Gibbs sampling, using Radford Neal's Algorithm 8 for non-conjugate priors \cite{Neal2000}. This is because the likelihood of $\+x$, which depends on the function $d$, is not a known conjugate-parametric form.\fxwarning{I'm not sure this is the right explanation.} We collapse $\+\pi$, but we do not collapse the other latent variables. Denote by $X$ the set of all $\+x_n$ and by $X_k = \left\{ \+x_n \in X : z_n = k \right\}$ and $N_k = | X_k |$. Then we derive the conditional posteriors by finding the Markov blankets of each node and using Bayes' rule. But because the likelihood of $\+x_n$ is note a standard parametric form, conditionals involving $X)k$ and $\+\mu_k$ cannot be sampled explicitly; we can only compute their likelihoods can be computed up to proportionality and resort to Metropolis-Hastings sampling.
\begin{align}\label{eq:conds}
p(z_n = k | \+\pi, \+x_n, \+\mu_k, \Lambda_k) &= p(z_n = k | \+\pi) \mathcal{N}(\+x_n | \+\mu_k, \Lambda_k) \label{eq:z}\\
p(\gamma_k | X_{z=k} , \beta_0, \gamma_0) &= \text{Gamma}\left( \gamma_k | \beta_0 + N_k, \gamma_0 + \sum_{\+x \in X_k} \+x \right) \label{eq:gamma}\\
p(\Lambda_k | X_k, \+\mu_k, \gamma_k, \Lambda_0, \nu) &\propto  p(X_k | \+\mu_k, \Lambda_k, \beta_k) \text{Wishart}(\Lambda_k | \Lambda_0, \nu) \label{eq:lambda}\\
p(\+\mu_k | X_k, \Lambda_k, \gamma_k, \+\mu_0, \kappa) &\propto p(X_k | \+\mu_k, \Lambda_k, \gamma_k) \mathcal{N}(\+\mu_k | \+\mu_0, (\kappa\Lambda)^{-1}) \label{eq:mu}
\end{align}
Note that $\+\pi$ remains collapsed; to evaluate \eqref{eq:z}, see \eqref{eq:mdp}.

\begin{algorithm}
\caption{Gibbs sample iteration}\label{alg:gibbs}
\begin{algorithmic}
\Require The current state of the Markov chain, $\Theta = \left\{ N, K, Z, \left\{ N_k, \+\mu_k, \Lambda_k, \gamma_k \right\}_{k=1}^K  \right\}$
\Ensure The distribution of the next state $\Theta^{*}$ preserves the Markov chain's invariant distribution.
\Statex

\For{$n := 1:N$}
  \If{$N_{z_n} = 1$}
    \State $\left( \+\mu_{K+1}, \Lambda_{K+1}, \gamma_{K+1} \right) := \left( \+\mu_{K+1}, \Lambda_{K+1}, \gamma_{K+1} \right)$ 
    \State Draw $\+\mu_j, \Lambda_j, \gamma_j$ from their priors (see section~\ref{sec:model}), for $j \in \{K+2, \ldots, K+m\}$
  \Else
    \State Draw $\+\mu_j, \Lambda_j, \gamma_j$ from their priors (see section~\ref{sec:model}), for $j \in \{K+1, \ldots, K+m\}$
  \EndIf
  \State Draw $z_n$ from 
  \[
  p(z_{n}=k|\+{z}_{-n},\+x_n,\+\mu_k,\Lambda_k,\gamma_k,\alpha)=\begin{cases}
  \frac{N_{k,-n}}{N-1+\alpha} p(\+x_n | \+\mu_k, \Lambda_k, \beta_k) & k < K\\
  \frac{\alpha / m}{N-1+\alpha} p(\+x_n | \+\mu_k, \Lambda_k, \beta_k) & \text{otherwise}
  \end{cases} \label{eq:nealalg8}
  \]
  \State Delete parameters for classes $k$ with $N_k = 0$ and relabel the $z_n$ to remove gaps.
\EndFor

\For{$k := 1:K$}
  \State Draw $\gamma_k$ using \eqref{eq:gamma}
  \State Metropolis-Hastings sample $\Lambda_k$ using likelihood \eqref{eq:lambda} and proposal $\text{Wishart}(\Lambda^{*} | p \Lambda, \nu)$ for some $p > 1$.
  \State Metropolis-Hastings sample $\+\mu_k$ using likelihood \eqref{eq:mu} and proposal $\mathcal{N}( \+\mu^{*} | C )$.
\EndFor

\end{algorithmic}
\end{algorithm}
